---
title: "QDR Metrics Report"
author: "Qualitative Data Repository"
date: "Generated `r Sys.time()`"
output:
  pdf_document: 
    includes:
      in_header: header.tex
  html_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE) 
opts <- options(knitr.kable.NA = "0")
library(httr)
library(readr)
library(tidyr)
library(dplyr)
library(ggplot2)
library(ggfortify)
library(stringi)
source("R/im.R")
source("R/deposits.R")
source("R/mdc.R")

### Configurable variables

# Custom start and end dates: use YYYY-MM-DD format
# These are specified from the workflow
# If these aren't provided, we'll: default to today's date as end date and to the beginning of the year as the start date
period_begin <- as.Date(Sys.getenv("period_begin"))
period_end <- as.Date(Sys.getenv("period_end"))

# Calculate dates: We use YTD if none are given
if (is.na(period_begin)) {
  period_begin <- as.Date(paste(as.integer(format(Sys.Date(), "%Y")),
                              "01-01", sep = "-"))}
if (is.na(period_end)){
  period_end <- Sys.Date()
}

# How many logins before we assume a user is QDR staff?
QDR_staff_logins <- 10


# Institutional members
members <-
  c(
    "American University",
    "Brown University",
    "Case Western",
    "Columbia University",
    "Cornell University",
    "Duke University",
    "Fordham",
    "George Mason",
    "George Washington University",
    "Marine Corps",
    "New York University",
    "Palliative Care Research Cooperative",
    "Princeton University",
    "Syracuse University",
    "University of California, Berkeley|UC Berkeley",
    "California, Merced|UC Merced",
    "University of California, San Francisco|UCSF",
    "University of Cincinnati",
    "University of Iowa",
    "University of Maryland Baltimore",
    "University of Michigan",
    "University of New Mexico",
    "University of North Carolina at Chapel Hill",
    "University of Texas at Austin",
    "Villanova University",
    "Virginia Tech",
    "West Virginia"
  )
# Add the prospective institutional member (input from github actions workflow)
extend <- Sys.getenv("extend_im")
if (extend!="") members <- c(members, extend)

# Set tokens and file location settings
dataverse_host <- "data.qdr.syr.edu"
dataverse_key <- Sys.getenv("DATAVERSE_TOKEN") # Github secret
drupal_folder <- paste0(Sys.getenv("EXTERNAL_GH_FOLDER"), "/drupal") # Github secret

# Drupal data, format as necessary
use <- read_csv(paste0(drupal_folder, "/use-stats.csv"))
login <- read_csv(paste0(drupal_folder, "/login-history.csv"))
login$Created <- as.Date(login$Created, format = "%m/%d/%Y")
login$`Last login` <- as.Date(login$`Last login`, format = "%a, %m/%d/%Y - %R")
login$Institution <- toupper(login$Institution)
login$Institution <- trimws(login$Institution)
login$full_name <- with(login, paste(`Last Name`, `First Name`, sep=", "))

# Use number of logins specified above for QDR membership
login$frequent_flyer <- login$`Logins during time period` > `QDR_staff_logins`
# This runs the helper function
login$member <- checkif_member(login, extend)
login$recent <- as.Date(login$`Last login`) > period_begin

# Calculate basic drupal summary stats
drupal_users <- nrow(login)
drupal_new_users <- sum(login$Created < period_end & login$Created > period_begin)
drupal_unique_logins <- as.integer(sum(login$recent))
drupal_logins_without_ff <- sum(login$`Logins during time period`[!login$frequent_flyer])

# Download and format dataverse data
dataverse <- paste0('http://', dataverse_host, "/api/info/metrics")
# Locally cache dataverse data as RDS files
# This is mainly useful because it reduces API calls when compiling to multiple formats (i.e. PDF+HTML)
if (file.exists("deposits.rds")) {
  deposits <- readRDS("deposits.rds")
  projects <- readRDS("projects.rds")
  mdc <- readRDS("mdc.rds")
} else {
  # This implements the search API to get all available data for all datasets
  deposits <- deposits_get(dataverse_host, dataverse_key)
  saveRDS(deposits, "deposits.rds")
  
  # Collect all depositor affiliations in the institutions variable
  institutions <- lapply(deposits$global_id, FUN=function(x, key) {
    # The only way to get institutional associations for depositors is to make a deposit-specific api query
    # So we query each individual deposit
    metadata <- content(GET(paste0("http://data.qdr.syr.edu",
                                   "/api/datasets/:persistentId/?persistentId=", 
                                   x), config=add_headers("X-Dataverse-Key"=key)))
    # Again, unnest the json into a tibble
    metadata <- 
      tibble(info=metadata$data$latestVersion$metadataBlocks$citation$fields) %>% 
      unnest_wider(info)
    
    affiliations <-
      metadata %>% filter(typeName == "author") %>% select("value") %>%
      unnest_auto(value) %>% unnest_auto(value)
    
    # check if there's no affiliation entry in the metadata
    if (is.null(affiliations$authorAffiliation)) {
      return(NA)
    }
    
    affiliations %>%
      select("authorAffiliation") %>% unnest_wider(authorAffiliation) %>%
      select("value") %>% .$value
  }, dataverse_key)
  
  # Run all the depositor affiliations against a regex of institutional members
  regex <- paste0("\\b(?:", paste(members, collapse="|"), ")\\b")
  matches <- lapply(institutions, function(x, regex) {
    matches <- stringr::str_extract(x, regex)
    dplyr::first(na.omit(matches))
  }, regex)
  deposits$member <- unlist(matches)
  
  saveRDS(deposits, "deposits.rds")
  
  # These are aggregate stats, only require one API call
  projects <- content(GET(paste0(dataverse, "/uniquedownloads/monthly")))
  projects <- bind_rows(projects$data)
  projects$date <- as.Date(paste0(projects$date, "-01"))
  saveRDS(projects, "projects.rds")
  
  mdc <- mdc_get(dataverse_host)
  saveRDS(mdc, "mdc.rds")
}
```

## Parameters

- Reporting period: `r period_begin` to `r period_end`.
- Non-member institutions considered: `r extend`

## Drupal

- QDR currently has `r drupal_users` users registered in Drupal. 
- `r drupal_new_users` of those users were created between `r period_begin` and `r period_end`.
- `r drupal_unique_logins` unique users have logged in since `r period_begin`.
- There have been at least `r drupal_unique_logins` unique logins since `r period_begin`. 

<!--Based on the time period used to generate the drupal login csv file:-->

<!--- If we subtract presumed QDR staff members (> `r QDR_staff_logins` logins/user), this number drops to `r drupal_logins_without_ff`. --> 

```{r drupal}
login %>% group_by(Institution) %>% summarise(users=length(`User status`)) %>% 
  arrange(desc(users)) %>% slice(1:10) %>%
  knitr::kable(row.names = FALSE, col.names = c("Institution", "Affiliated users"),
               caption="Top institutional affiliations by number of users")

use %>% select(-`Views today`) %>% arrange(desc(`Total views`)) %>% 
  slice(1:10) %>%
  mutate(Title = paste0("[", Title, "](", 
                      "https://qdr.syr.edu", `Link to Content`, ")")) %>%
  select(!`Link to Content`) %>%
  knitr::kable(row.names = FALSE, col.names = c("Title", "Views"),
               caption="Most viewed QDR pages")
```

```{r members_drupal}
members_drupal <- count(group_by(login, member))

members_drupal <- login %>% group_by(member) %>% summarize(recent=sum(recent)) %>% full_join(members_drupal)

im_drupal <- login %>% group_by(member) %>% filter(Created < period_end & Created > period_begin) %>% 
  summarize(created_since=length(`Name/Email`)) %>% full_join(members_drupal) %>%
  arrange(desc(n)) %>% replace_na(list(created_since=0))

# Here we write drupal member data to an external file
# This was by request, in order to have an easier parsed version than the table
im_drupal %>% filter(!is.na(member)) %>%
  replace_na(replace = list(created_since=0, recent=FALSE, n=0)) %>%
  write_csv(paste0(Sys.getenv("EXTERNAL_GH_FOLDER"), "/im_drupal.csv"))

# Now display the same data as a table in the PDF/HTML
im_drupal %>%
  knitr::kable(col.names = c("Institution", 
                             paste("Created between", period_begin, "and", period_end),
                             paste("Min unique logins since", period_begin),
                             "Total users"),
               caption="QDR Accounts and Logins by Institutional Members")
```

## Dataverse

```{r total_projects}
dataverse_total <- nrow(deposits)

deposits_new <- deposits %>% filter(as.Date(createdAt) < period_end & as.Date(createdAt) > period_begin) %>%
  summarize(n=n()) %>% .$n

published_new <- deposits %>% filter(as.Date(published_at) < period_end & as.Date(published_at) > period_begin) %>%
  summarize(n=n()) %>% .$n
```

- There are currently `r dataverse_total` projects. 
- `r deposits_new` new deposits were initiated between `r period_begin` and `r period_end`.
- `r published_new` projects were published between `r period_begin` and `r period_end`.

```{r subject}
deposits %>% group_by(versionState) %>% summarize(n=n()) %>%
  knitr::kable(caption="Total deposits by publication status", col.names=c("Status", "Deposits"))

deposits_subj <- deposits %>% 
  filter(versionState=="Published") %>% select(c("subjects")) %>%
  unnest_longer(subjects) %>% group_by(subjects) %>% summarize(published=n())

  
if  (nrow(subset(deposits, as.Date(createdAt) < period_end &
               as.Date(createdAt) > period_begin))==0) {
  
  deposits_subj <- deposits %>% 
    filter(as.Date(published_at) < period_end & as.Date(published_at) > period_begin & 
             versionState=="Published") %>%
    select(c("subjects")) %>% unnest_longer(subjects) %>% group_by(subjects) %>%
    summarize(period=n()) %>% full_join(deposits_subj)

  } else {
  deposits_subj$period <- 0
}

# Any unpublished projects?
if (any(deposits$versionState=="Unpublished")) {
  deposits_subj <- deposits %>% filter(versionState=="Unpublished") %>% select(c("subjects")) %>%
    unnest_longer(subjects) %>% group_by(subjects) %>%
    summarize(unpublished=n()) %>% full_join(deposits_subj)
} else {
  deposits_subj$unpublished <- 0
}

# Any unpublished in the time period specified?
if (any(subset(deposits, as.Date(createdAt) < period_end &
               as.Date(createdAt) > period_begin)$versionState=="Unpublished")) {
  deposits_subj <- deposits %>% filter(versionState == "Unpublished" & 
                                      as.Date(createdAt) < period_end &
                                      as.Date(createdAt) > period_begin) %>%
    select(c("subjects")) %>% unnest_longer(subjects) %>% group_by(subjects) %>%
    summarize(unpublished_period=n()) %>% full_join(deposits_subj)
} else {
  deposits_subj$unpublished_period <- 0
}

deposits_subj[, c("subjects", "published", 
                  "unpublished", "period", "unpublished_period")] %>% 
  knitr::kable(caption="Deposits by subject",
               col.names=c("Subject", "Total published", "Total unpublished",
                           "Published in reporting period", 
                           "Unpublished in reporting period"))
```

```{r top}
projects %>% group_by(pid) %>% summarize(sum(count)) %>% 
  arrange(desc(`sum(count)`)) %>% slice(1:5) %>%
  mutate(pid = paste0("[", deposits$name[deposits$global_id %in% pid], "](", 
                      "https://data.qdr.syr.edu/dataset.xhtml?persistentId=", 
                      pid, ")")) %>%
  knitr::kable(caption="Top 5 all time downloaded projects",
             row.names = FALSE, col.names = c("Project", "Downloads"))

projects %>% group_by(pid) %>% filter(date < period_end &
                      date > period_begin) %>% 
  summarize(sum(count)) %>% arrange(desc(`sum(count)`)) %>% slice(1:5) %>%
  mutate(pid = paste0("[", deposits$name[deposits$global_id %in% pid], "](", 
                      "https://data.qdr.syr.edu/dataset.xhtml?persistentId=", 
                      pid, ")")) %>%
  knitr::kable(caption=paste("Top 5 downloaded projects between ", period_begin,
                             "and", period_end),
             row.names = FALSE, col.names = c("Project", "Downloads"))
```


### Institutional members

```{r member_projects}
member_deposits <- deposits %>% filter(!is.na(member)) %>% 
  group_by(member, versionState) %>% 
  tally()

im_deposit <- 
  deposits %>% filter(!is.na(member) & as.Date(createdAt) < period_end & as.Date(createdAt) > period_begin) %>% 
  group_by(member, versionState) %>% 
  summarize(period=length(global_id)) %>% full_join(member_deposits) %>%
  pivot_wider(names_from=versionState, names_sep=".", values_from=c(n, period), values_fill=0) 

im_deposit %>% knitr::kable(caption="Deposits associated with institutional members",
               col.names = c("Institutional member name", 
                             "Total published", "Total unpublished",
                             "Published in reporting period", "Unpublished in reporting period"))

# Write the institutional member dataverse data to an external file as well
# Same as the table included immediately above
# This is easier to parse than the table that is included in the PDF/HTML
im_deposit %>% 
  replace_na(list(n.Published=0, n.Unpublished=0, period.Published=0, period.Unpublished=0)) %>%
  write_csv(paste0(Sys.getenv("EXTERNAL_GH_FOLDER"), "/im_deposits.csv"))

projects %>% group_by(global_id=pid) %>% summarize(downloads=sum(count)) %>%
  inner_join(deposits) %>% select("name", "global_id", "member", "downloads") %>%
  filter(!is.na(member)) %>% arrange(desc(downloads)) %>%
  group_by(member) %>% slice_head(n=3) %>%
  mutate(name = paste0("[", name, "](", 
                      "https://data.qdr.syr.edu/dataset.xhtml?persistentId=", 
                      global_id, ")")) %>% select(!global_id) %>%
  knitr::kable(caption="Top 3 downloads from each member institution",
               col.names = c("Deposit", "Member institution", "Downloads"),
               linesep = "\\addlinespace")
```

## Make Data Count (MDC)

### Understanding MDC data

MDC reports statistics as "total" and "unique". 

- Unique" views/downloads correspond to the number of unique sessions in which a data project has been viewed/downloaded. Roughly translates to: "X people have viewed/downloaded this project or its files."
- Total views/downloads correspond to the number of times a project has been viewed/downloaded, including, e.g., multiple counts for multiple views/downloads in the same session or by the same user.


```{r mdc}
if (exists("mdc")) {
  stats_ts <- ts(mdc[, -1], frequency=12, start=strsplit(mdc$date[1], split="-")[[1]])
} else {
  stats_ts <- ts(downloads[, -1], frequency=12, 
                start=strsplit(downloads$date[1], split="-")[[1]])
}

autoplot(stats_ts, facets = FALSE) +  labs(title="Statistics", color="Statistic") +
    scale_y_continuous(labels=scales::comma)

```

