---
title: "Metrics report"
author: "Qualitative Data Repository"
date: "Generated `r Sys.time()`"
output:
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE) 
opts <- options(knitr.kable.NA = "")
library(httr)
library(readr)
library(tidyr)
library(dplyr)
library(ggplot2)
library(ggfortify)
source("helper.R")

# Set tokens and file location settings
dataverse_host <- "http://data.qdr.syr.edu"
dataverse_key <- Sys.getenv("DATAVERSE_TOKEN")
drupal_folder <- paste0(Sys.getenv("EXTERNAL_GH_FOLDER"), "/drupal")
drupal_use <- paste0(drupal_folder, "/use-stats.csv")
drupal_login <- paste0(drupal_folder, "/login-history.csv")

# Calculate dates
period_begin <- as.Date(paste(as.integer(format(Sys.Date(), "%Y"))-1,
                              "01-01", sep = "-"))
period_end <- as.Date(paste(as.integer(format(Sys.Date(), "%Y"))-1,
                              "12-31", sep = "-"))
# Drupal data, format as necessary
use <- read_csv(drupal_use)
login <- read_csv(drupal_login)

#login <- login[complete.cases(login), ]
login$Created <- as.Date(login$Created, format = "%m/%d/%Y")
login$`Last login` <- as.Date(login$`Last login`, format = "%a, %m/%d/%Y - %R")
login$Institution <- toupper(login$Institution)
login$Institution <- trimws(login$Institution)
login$full_name <- with(login, paste(`Last Name`, `First Name`, sep=", "))

# This runs the helper function
extend_members <- paste0(Sys.getenv("EXTERNAL_GH_FOLDER"), "/extend_members.csv")
if (file.exists(extend_members)) {
  extend_members <- read_csv(extend_members)
  extension_names <- paste(extend_members$name, collapse=", ")
}
login$member <- checkif_member(login, extend_members)

# Basic statistics
drupal_users <- nrow(login)
ff_cutoff <- 10 # How many logins before a user is considered a "frequent flyer"?
login$frequent_flyer <- login$`Logins during time period` > ff_cutoff
drupal_new_users <- sum(login$Created < period_end & login$Created > period_begin)

drupal_all_logins <- sum(login$`Logins during time period`)
drupal_unique_logins <- sum(login$`Last login` > period_begin)
drupal_logins_without_ff <- sum(login$`Logins during time period`[!login$frequent_flyer])

# Dataverse data
dataverse <- paste0(dataverse_host, "/api/info/metrics")
# This implements the search API to get all available data for all datasets
deposits <- content(GET(paste0(dataverse_host,
                               "/api/search/?q=*&type=dataset&per_page=1000"),
                        config=add_headers("X-Dataverse-Key"=dataverse_key)))
deposits <- tibble(deposit=deposits$data$items)
deposits <- deposits %>% unnest_wider(deposit)
deposits$name <- stringr::str_trunc(deposits$name, 75, "right")
deposits$versionState <- recode(deposits$versionState, RELEASED="Published", DRAFT="Unpublished")
```

## Parameters

- Special period: `r period_begin` to `r period_end`.
- Extensions to institutional member list: `r extension_names`

## Drupal

- QDR currently has `r drupal_users` users registered in Drupal. 
- `r drupal_new_users` of those users were created between `r period_begin` and `r period_end`.
- `r drupal_unique_logins` unique users have logged in since `r period_begin`.

Based on the time period used to generate the `r drupal_login` file:

- There have been `r drupal_all_logins` logins. 
- If we subtract the "frequent flyers" (`r ff_cutoff` logins/user), this number drops to `r drupal_logins_without_ff`.  

```{r drupal}
login %>% group_by(Institution) %>% summarise(users=length(`User status`)) %>% 
  arrange(desc(users)) %>% slice(1:10) %>%
  knitr::kable(row.names = FALSE, col.names = c("Institution", "Affiliated users"),
               caption="Top institutional affiliations by number of users")

use %>% select(-`Views today`) %>% arrange(desc(`Total views`)) %>% 
  slice(1:10) %>%
  mutate(Title = paste0("[", Title, "](", 
                      "https://qdr.syr.edu", `Link to Content`, ")")) %>%
  select(!`Link to Content`) %>%
  knitr::kable(row.names = FALSE, col.names = c("Title", "Views"),
               caption="Most accessed drupal pages")
```

```{r members_drupal}
members_drupal <- count(group_by(login, member))

members_drupal <- login %>% group_by(member) %>% summarize(logins=sum(`Logins during time period`)) %>% full_join(members_drupal)

login %>% group_by(member) %>% filter(Created < period_end & Created > period_begin) %>% 
  summarize(created_since=length(`Name/Email`)) %>% full_join(members_drupal) %>%
  arrange(desc(n)) %>% # replace_na(list(created_since=0)) %>%
  knitr::kable(col.names = c("Institution", paste("Created between", 
                                                  period_begin, "and", period_end),
                             "Logins", "Total users"),
               caption="User statistics of institutional member drupal accounts")
```

## Dataverse

```{r total_projects}
dataverse_total <- nrow(deposits)
deposits_new <- sum(as.Date(deposits$createdAt) < period_end & as.Date(deposits$createdAt) > period_begin)
```

- There are currently `r dataverse_total` projects. 
- `r deposits_new` are from the period between `r period_begin` and `r period_end`.

```{r subject}
deposits$versionState %>% table() %>% 
  knitr::kable(caption="Deposits by publication status", col.names=c("Status", "Deposits"))

subject <- content(GET(paste0(dataverse, "/datasets/bySubject")))
subject <- bind_rows(subject$data)
subject$count <- round(100*(subject$count/sum(subject$count)), digits=2)
knitr::kable(subject, col.names=c("Subject", "Percentage"), 
             caption="Distribution of deposit subjects")
```

```{r top}
projects <- content(GET(paste0(dataverse, "/uniquedownloads/monthly")))
projects <- bind_rows(projects$data)
projects$date <- as.Date(paste0(projects$date, "-01"))

projects %>% group_by(pid) %>% summarize(sum(count)) %>% 
  arrange(desc(`sum(count)`)) %>% slice(1:5) %>%
  mutate(pid = paste0("[", deposits$name[deposits$global_id %in% pid], "](", 
                      "https://data.qdr.syr.edu/dataset.xhtml?persistentId=", 
                      pid, ")")) %>%
  knitr::kable(caption="Top 5 all time downloaded projects",
             row.names = FALSE, col.names = c("Project", "Downloads"))

projects %>% group_by(pid) %>% filter(date < period_end) %>% 
  summarize(sum(count)) %>% arrange(desc(`sum(count)`)) %>% slice(1:5) %>%
  mutate(pid = paste0("[", deposits$name[deposits$global_id %in% pid], "](", 
                      "https://data.qdr.syr.edu/dataset.xhtml?persistentId=", 
                      pid, ")")) %>%
  knitr::kable(caption=paste("Top 5 downloaded projects since", period_end),
             row.names = FALSE, col.names = c("Project", "Downloads"))
```


### Institutional members

```{r member_projects}
# Can we detect institutional members from our routine earlier based on names?
deposit_members <- lapply(deposits$contacts, function(x, login) {
  names <- tibble(x=x) %>% hoist(x, "name") %>% .$name
  match <- str_which(login$full_name, paste(names, collapse="|"))[1]
  if (length(match)==0) return(NA)
  return(login$member[match])
}, login)
deposits <- tibble(member=deposit_members) %>% unnest(member) %>% bind_cols(deposits)

member_deposits <- deposits %>% filter(!is.na(member)) %>% 
  group_by(member, versionState) %>% 
  tally()
#  knitr::kable(caption = "Deposits associated with member", 
#               col.names = c("Institution", "Status", "Deposits"))

deposits %>% filter(!is.na(member) & as.Date(createdAt) > period_end) %>% 
  group_by(member, versionState) %>% 
  summarize(period=length(global_id)) %>% full_join(member_deposits) %>%
  knitr::kable(caption="Deposits associated with member",
               col.names = c("Institution", "Status", 
                             paste(period_begin, "to", period_end),
                             "Total"))

projects %>% group_by(global_id=pid) %>% summarize(downloads=sum(count)) %>%
  inner_join(deposits) %>% select("name", "global_id", "member", "downloads") %>%
  filter(!is.na(member)) %>% arrange(desc(downloads)) %>%
  group_by(member) %>% slice_head(n=3) %>%
  mutate(name = paste0("[", name, "](", 
                      "https://data.qdr.syr.edu/dataset.xhtml?persistentId=", 
                      global_id, ")")) %>% select(!global_id) %>%
  knitr::kable(caption="Top 3 downloads from each member institution",
               col.names = c("Deposit", "Member institution", "Downloads"),
               linesep = "\\addlinespace")
```

## Make Data Count
```{r views}
views_total <- content(GET(paste0(dataverse, '/makeDataCount/viewsTotal/monthly')))
views_total <- bind_rows(views_total$data)
names(views_total) <- c("date", "total")

views_unique <- content(GET(paste0(dataverse, '/makeDataCount/viewsUnique/monthly')))
views_unique <- bind_rows(views_unique$data)
names(views_unique) <- c("date", "unique")

both <- merge(views_total, views_unique, all=TRUE, by="date")
both_ts <- ts(both[, -1], frequency=12, 
              start=strsplit(both$date[1], split="-")[[1]])
autoplot(both_ts, facets = FALSE) + labs(title="Views", color="Statistic") 
```

```{r downloads}
downloads_total <- content(GET(paste0(dataverse, '/makeDataCount/downloadsTotal/monthly')))
downloads_total <- bind_rows(downloads_total$data)
names(downloads_total) <- c("date", "total")

downloads_unique <- content(GET(paste0(dataverse, '/makeDataCount/downloadsUnique/monthly')))
downloads_unique <- bind_rows(downloads_unique$data)
names(downloads_unique) <- c("date", "unique")

both <- merge(downloads_total, downloads_unique, all=TRUE, by="date")
both_ts <- ts(both[, -1], frequency=12, 
              start=strsplit(both$date[1], split="-")[[1]])
autoplot(both_ts, facets = FALSE) +  labs(title="Downloads", color="Statistic") 
```

