---
title: "Metrics report"
author: "Qualitative Data Repository"
date: "Generated `r Sys.time()`"
output:
  pdf_document: 
    includes:
      in_header: header.tex
  html_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE) 
opts <- options(knitr.kable.NA = "")
library(httr)
library(readr)
library(tidyr)
library(dplyr)
library(ggplot2)
library(ggfortify)
library(stringi)
source("helper.R")

### Configurable variables

# Custom start and end dates: use YYYY-MM-DD format
# If these aren't provided, we'll: default to today's date as end date and to the beginning of the year as the start date
period_begin <- NULL
period_end <- NULL

# How many logins before we assume a user is QDR staff?
QDR_staff_logins <- 10


# Institutional members
members <-
  c(
    "American University",
    "Brown University",
    "Case Western",
    "Columbia University",
    "Cornell University",
    "Duke University",
    "Fordham",
    "George Mason",
    "George Washington University",
    "Macalester",
    "Marine Corps",
    "3ie",
    "New York University",
    "Palliative Care Research Cooperative",
    "Princeton University",
    "Syracuse University",
    "University of California, Berkeley|UC Berkeley",
    "California, Merced",
    "University of Cincinnati",
    "University of Iowa",
    "University of Maryland Baltimore",
    "University of Michigan",
    "University of New Mexico",
    "University of North Carolina at Chapel Hill",
    "University of Texas at Austin",
    "Villanova University",
    "Virginia Tech",
    "West Virginia"
  )
# Add the prospective institutional member
extend <- Sys.getenv("extend_im")
if (extend!="") members <- c(members, extend)


# Set tokens and file location settings
dataverse_host <- "http://data.qdr.syr.edu"
dataverse_key <- Sys.getenv("DATAVERSE_TOKEN")
drupal_folder <- paste0(Sys.getenv("EXTERNAL_GH_FOLDER"), "/drupal")

# Calculate dates: We use YTD if none are given
if (is.null(period_begin)) {
  period_begin <- as.Date(paste(as.integer(format(Sys.Date(), "%Y")),
                              "01-01", sep = "-"))}
if (is.null(period_end)){
  period_end <- Sys.Date()
}

# Drupal data, format as necessary
use <- read_csv(paste0(drupal_folder, "/use-stats.csv"))
login <- read_csv(paste0(drupal_folder, "/login-history.csv"))

#login <- login[complete.cases(login), ]
login$Created <- as.Date(login$Created, format = "%m/%d/%Y")
login$`Last login` <- as.Date(login$`Last login`, format = "%a, %m/%d/%Y - %R")
login$Institution <- toupper(login$Institution)
login$Institution <- trimws(login$Institution)
login$full_name <- with(login, paste(`Last Name`, `First Name`, sep=", "))

# Use number of logins specified above for QDR membership
login$frequent_flyer <- login$`Logins during time period` > QDR_staff_logins
# This runs the helper function
login$member <- checkif_member(login, extend) # Needs the extension too!

# Basic statistics
drupal_users <- nrow(login)

drupal_new_users <- sum(login$Created < period_end & login$Created > period_begin)

drupal_all_logins <- sum(login$`Logins during time period`)
drupal_unique_logins <- sum(login$`Last login` > period_begin)
drupal_logins_without_ff <- sum(login$`Logins during time period`[!login$frequent_flyer])

# Dataverse data
dataverse <- paste0(dataverse_host, "/api/info/metrics")
# This implements the search API to get all available data for all datasets
deposits <- content(GET(paste0(dataverse_host,
                               "/api/search/?q=*&type=dataset&per_page=1000"),
                        config=add_headers("X-Dataverse-Key"=dataverse_key)))
deposits <- tibble(deposit=deposits$data$items)
deposits <- deposits %>% unnest_wider(deposit)
 
# Check for institutional members
institutions <- lapply(deposits$global_id, FUN=function(x, key) {
  metadata <- content(GET(paste0("http://data.qdr.syr.edu",
         "/api/datasets/:persistentId/?persistentId=", 
         x), config=add_headers("X-Dataverse-Key"=key)))
  
  metadata <- 
    tibble(info=metadata$data$latestVersion$metadataBlocks$citation$fields) %>% 
    unnest_wider(info)
  
  affiliations <-
    metadata %>% filter(typeName == "author") %>% select("value") %>%
    unnest_auto(value) %>% unnest_auto(value)
  
  # check if there's no affiliation entry in the metadata
  if (is.null(affiliations$authorAffiliation)) {
    return(NA)
  }
  
  affiliations %>%
    select("authorAffiliation") %>% unnest_wider(authorAffiliation) %>%
    select("value") %>% .$value
}, dataverse_key)

regex <- paste0("\\b(?:", paste(members, collapse="|"), ")\\b")
matches <- lapply(institutions, function(x, regex) {
  matches <- stringr::str_extract(x, regex)
  dplyr::first(na.omit(matches))
}, regex)
deposits$member <- unlist(matches)

deposits$name <- stringr::str_trunc(deposits$name, 75, "right")
deposits$versionState <- recode(deposits$versionState, RELEASED="Published", DRAFT="Unpublished")
```

## Parameters

- Reporting period: `r period_begin` to `r period_end`.
- Non-member institutions considered: `r extend`

## Drupal

- QDR currently has `r drupal_users` users registered in Drupal. 
- `r drupal_new_users` of those users were created between `r period_begin` and `r period_end`.
- `r drupal_unique_logins` unique users have logged in since `r period_begin`.

Based on the time period used to generate the drupal login csv file:

- There have been `r drupal_all_logins` logins. 
- If we subtract presumed QDR staff members (>`QDR_staff_logins` logins/user), this number drops to `r drupal_logins_without_ff`.  

```{r drupal}
login %>% group_by(Institution) %>% summarise(users=length(`User status`)) %>% 
  arrange(desc(users)) %>% slice(1:10) %>%
  knitr::kable(row.names = FALSE, col.names = c("Institution", "Affiliated users"),
               caption="Top institutional affiliations by number of users")

use %>% select(-`Views today`) %>% arrange(desc(`Total views`)) %>% 
  slice(1:10) %>%
  mutate(Title = paste0("[", Title, "](", 
                      "https://qdr.syr.edu", `Link to Content`, ")")) %>%
  select(!`Link to Content`) %>%
  knitr::kable(row.names = FALSE, col.names = c("Title", "Views"),
               caption="Most viewed QDR pages")
```

```{r members_drupal}
members_drupal <- count(group_by(login, member))

members_drupal <- login %>% group_by(member) %>% summarize(logins=sum(`Logins during time period`)) %>% full_join(members_drupal)

login %>% group_by(member) %>% filter(Created < period_end & Created > period_begin) %>% 
  summarize(created_since=length(`Name/Email`)) %>% full_join(members_drupal) %>%
  arrange(desc(n)) %>% # replace_na(list(created_since=0)) %>%
  knitr::kable(col.names = c("Institution", paste("Created between", 
                                                  period_begin, "and", period_end),
                             "Logins", "Total users"),
               caption="QDR Accounts and Logins by Institutional Members")
```

## Dataverse

```{r total_projects}
dataverse_total <- nrow(deposits)
deposits_published <- deposits %>% filter(versionState == "Published")
deposits_unpublished <- deposits %>% filter(versionState == "Unpublished")
deposits_new <- sum(as.Date(deposits$createdAt) < period_end & as.Date(deposits$createdAt) > period_begin)
published_new <- sum(as.Date(deposits_published$published_at) < period_end & as.Date(deposits_published$published_at) > period_begin)
```

- There are currently `r dataverse_total` projects. 
- `r deposits_new` new deposits were initiated between `r period_begin` and `r period_end`.
- `r published_new` projects were published between `r period_begin` and `r period_end`.

```{r subject}
deposits %>% group_by(versionState) %>% summarize(n=n()) %>%
  knitr::kable(caption="Total deposits by publication status", col.names=c("Status", "Deposits"))

deposits_published %>% select(c("subjects")) %>%
  unnest_longer(subjects) %>% group_by(subjects) %>%
  summarize(pct=round(100*(n()/nrow(deposits_published)), digits=2)) %>%
  knitr::kable(caption="Published deposits by subject (projects can have multiple subjects)", 
               col.names=c("Subject", "Percentage of projects")) 

deposits_unpublished %>% select(c("subjects")) %>%
  unnest_longer(subjects) %>% group_by(subjects) %>%
  summarize(pct=round(100*(n()/nrow(deposits_unpublished)), digits=2)) %>%
  knitr::kable(caption="Draft deposits by subject (projects can have multiple subjects)", 
               col.names=c("Subject", "Percentage of projects")) 

deposits_published %>% filter(as.Date(createdAt) < period_end &
                      as.Date(createdAt) > period_begin) %>%
  select(c("subjects")) %>%
  unnest_longer(subjects) %>% group_by(subjects) %>%
  summarize(pct=round(100*(n()/nrow(filter(deposits_published, as.Date(createdAt) < period_end &
                      as.Date(createdAt) > period_begin))), digits=2)) %>%
  knitr::kable(caption=paste("Deposits by subject published between", period_begin,
                             "and", period_end),
               col.names=c("Subject", "Percentage of projects"))


deposits_unpublished %>% filter(as.Date(createdAt) < period_end &
                      as.Date(createdAt) > period_begin) %>%
  select(c("subjects")) %>%
  unnest_longer(subjects) %>% group_by(subjects) %>%
  summarize(pct=round(100*(n()/nrow(filter(deposits_unpublished, as.Date(createdAt) < period_end &
                      as.Date(createdAt) > period_begin))), digits=2)) %>%
  knitr::kable(caption=paste("Draf deposits by subject created between", period_begin,
                             "and", period_end),
               col.names=c("Subject", "Percentage of projects"))
```

```{r top}
projects <- content(GET(paste0(dataverse, "/uniquedownloads/monthly")))
projects <- bind_rows(projects$data)
projects$date <- as.Date(paste0(projects$date, "-01"))

projects %>% group_by(pid) %>% summarize(sum(count)) %>% 
  arrange(desc(`sum(count)`)) %>% slice(1:5) %>%
  mutate(pid = paste0("[", deposits$name[deposits$global_id %in% pid], "](", 
                      "https://data.qdr.syr.edu/dataset.xhtml?persistentId=", 
                      pid, ")")) %>%
  knitr::kable(caption="Top 5 all time downloaded projects",
             row.names = FALSE, col.names = c("Project", "Downloads"))

projects %>% group_by(pid) %>% filter(date < period_end) %>% 
  summarize(sum(count)) %>% arrange(desc(`sum(count)`)) %>% slice(1:5) %>%
  mutate(pid = paste0("[", deposits$name[deposits$global_id %in% pid], "](", 
                      "https://data.qdr.syr.edu/dataset.xhtml?persistentId=", 
                      pid, ")")) %>%
  knitr::kable(caption=paste("Top 5 downloaded projects between ", period_begin,
                             "and", period_end),
             row.names = FALSE, col.names = c("Project", "Downloads"))
```


### Institutional members

```{r member_projects}
member_deposits <- deposits %>% filter(!is.na(member)) %>% 
  group_by(member, versionState) %>% 
  tally()
#  knitr::kable(caption = "Deposits associated with member", 
#               col.names = c("Institution", "Status", "Deposits"))

deposits %>% filter(!is.na(member) & as.Date(createdAt) > period_end) %>% 
  group_by(member, versionState) %>% 
  summarize(period=length(global_id)) %>% full_join(member_deposits) %>%
  knitr::kable(caption="Deposits associated with institutional members",
               col.names = c("Institution", "Status", 
                             "Reporting period",
                             "Total"))

projects %>% group_by(global_id=pid) %>% summarize(downloads=sum(count)) %>%
  inner_join(deposits) %>% select("name", "global_id", "member", "downloads") %>%
  filter(!is.na(member)) %>% arrange(desc(downloads)) %>%
  group_by(member) %>% slice_head(n=3) %>%
  mutate(name = paste0("[", name, "](", 
                      "https://data.qdr.syr.edu/dataset.xhtml?persistentId=", 
                      global_id, ")")) %>% select(!global_id) %>%
  knitr::kable(caption="Top 3 downloads from each member institution",
               col.names = c("Deposit", "Member institution", "Downloads"),
               linesep = "\\addlinespace")
```

## Make Data Count (MDC)

### Understanding MDC data

MDC reports statistics as "total" and "unique". 
- Unique" views/downloads correspond to the number of unique sessions in which a data project has been viewed/downloaded. Roughly translates to: "X people have viewed/downloaded this project or its files."
- Total views/downloads correspond to the number of times a project has been viewed/downloaded, including, e.g., multiple counts for multiple views/downloads in the same session or by the same user.

```{r views}
views_total <- content(GET(paste0(dataverse, '/makeDataCount/viewsTotal/monthly')))
views_total <- bind_rows(views_total$data)
names(views_total) <- c("date", "total")

views_unique <- content(GET(paste0(dataverse, '/makeDataCount/viewsUnique/monthly')))
views_unique <- bind_rows(views_unique$data)
names(views_unique) <- c("date", "unique")

both <- merge(views_total, views_unique, all=TRUE, by="date")
both_ts <- ts(both[, -1], frequency=12, 
              start=strsplit(both$date[1], split="-")[[1]])
autoplot(both_ts, facets = FALSE) + labs(title="MDC Views", color="Statistic") 
```



```{r downloads}
downloads_total <- content(GET(paste0(dataverse, '/makeDataCount/downloadsTotal/monthly')))
downloads_total <- bind_rows(downloads_total$data)
names(downloads_total) <- c("date", "total")

downloads_unique <- content(GET(paste0(dataverse, '/makeDataCount/downloadsUnique/monthly')))
downloads_unique <- bind_rows(downloads_unique$data)
names(downloads_unique) <- c("date", "unique")

both <- merge(downloads_total, downloads_unique, all=TRUE, by="date")
both_ts <- ts(both[, -1], frequency=12, 
              start=strsplit(both$date[1], split="-")[[1]])
autoplot(both_ts, facets = FALSE) +  labs(title="MDC Downloads", color="Statistic") 
```

