---
title: "Metrics"
author: "Qualitative Data Repository"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE) 
library(httr)
library(readr)
library(tidyr)
library(dplyr)
library(ggplot2)
library(ggfortify)
source("helper.R")

dataverse_host <- "http://data.qdr.syr.edu"
dataverse_key <- Sys.getenv("DATAVERSE_TOKEN")
drupal_use <- "drupal-metrics/use-stats.csv"
login_file <- "drupal-metrics/login-history-sample.csv"

# Calculate dates
today <- Sys.Date()
period_end <- as.Date("2021-01-01")

# Drupal data, format as necessary
use <- read_csv(drupal_use)
login <- read_csv(login_file)

#login <- login[complete.cases(login), ]
login$Created <- as.Date(login$Created, format = "%m/%d/%Y")
login$`Last login` <- as.Date(login$`Last login`, format = "%a, %m/%d/%Y - %R")
login$Institution <- toupper(login$Institution)
login$Institution <- trimws(login$Institution)
login$full_name <- with(login, paste(`Last Name`, `First Name`, sep=", "))

login$member <- checkif_member(login)
#institutional_members <- login[login$member, ]
#institutional_usercount <- sum(login$member)

drupal_users <- nrow(login)
login$frequent_flyer <- login$`Logins during time period` > 10
drupal_new_users <- sum(login$Created > period_end)

drupal_all_logins <- sum(login$`Logins during time period`)
drupal_unique_logins <- sum(login$`Last login` > period_end)
drupal_logins_without_ff <- sum(login$`Logins during time period`[!login$frequent_flyer])

# Dataverse data
dataverse <- paste0(dataverse_host, "/api/info/metrics")
# This implements the search API to get all available data for all datasets
deposits <- content(GET(paste0(dataverse_host,
                               "/api/search/?q=*&type=dataset&per_page=1000"),
                        config=add_headers("X-Dataverse-Key"=dataverse_key)))
deposits <- tibble(deposit=deposits$data$items)
deposits <- deposits %>% unnest_wider(deposit)
```

## Parameters

- Arbitrarily defined end-point date is set to `r period_end`.

## Drupal

- QDR currently has `r drupal_users` users registered in Drupal. 
- `r drupal_new_users` of those users were created since `r period_end`. 
- `r drupal_unique_logins` unique users have logged in since `r period_end`.

Based on the time period used to generate the `r login_file` file:

- There have been `r drupal_all_logins` logins. 
- If we subtract the "frequent flyers" (arbitrary number of login/user), this number drops to `r drupal_logins_without_ff`.  

```{r drupal}
login %>% group_by(Institution) %>% summarise(users=length(`User status`)) %>% 
  arrange(desc(users)) %>% slice(1:10) %>%
  knitr::kable(row.names = FALSE, col.names = c("Institution", "Affiliated users"),
               caption="Top institutional affiliations by number of users")

use %>% select(-`Views today`) %>% arrange(desc(`Total views`)) %>% slice(1:10) %>%
  knitr::kable(row.names = FALSE, col.names = c("Title", "Views", "Link"),
               caption="Most accessed drupal pages")
```

```{r members_drupal}
members_drupal <- count(group_by(login, member))

members_drupal <- login %>% group_by(member) %>% summarize(logins=sum(`Logins during time period`)) %>% full_join(members_drupal)

login %>% group_by(member) %>% filter(Created > period_end) %>% 
  summarize(created_since=length(`Name/Email`)) %>% full_join(members_drupal) %>%
  arrange(desc(n)) %>% replace_na(list(created_since=0)) %>%
  knitr::kable(col.names = c("Institution", paste("Created since", period_end),
                             "Logins", "Total users"),
               caption="Statistics of institutional member accounts")
```

## Dataverse

```{r total_projects}
dataverse_total <- nrow(deposits)
deposits_new <- sum(as.Date(deposits$createdAt) > period_end)
```

- There are currently `r dataverse_total` projects. 
- `r deposits_new` are from the period beginning `r period_end`.

```{r subject}
deposits$versionState %>% table() %>% 
  knitr::kable(caption="Deposit publication status", col.names=c("Status", "Deposits"))

subject <- content(GET(paste0(dataverse, "/datasets/bySubject")))
subject <- bind_rows(subject$data)
subject$count <- round(100*(subject$count/sum(subject$count)), digits=2)
knitr::kable(subject, col.names=c("Subject", "Percentage"), 
             caption="Distribution of deposit subjects")
```

```{r top}
projects <- content(GET(paste0(dataverse, "/uniquedownloads/monthly")))
projects <- bind_rows(projects$data)
projects$date <- as.Date(paste0(projects$date, "-01"))

projects %>% group_by(pid) %>% summarize(sum(count)) %>% 
  arrange(desc(`sum(count)`)) %>% slice(1:5) %>%
  knitr::kable(caption="Top 5 all time downloaded projects",
             row.names = FALSE, col.names = c("Persistent ID", "Downloads"))

projects %>% group_by(pid) %>% filter(date > period_end) %>% 
  summarize(sum(count)) %>% arrange(desc(`sum(count)`)) %>% slice(1:5) %>%
  knitr::kable(caption=paste("Top 5 downloaded projects since", period_end),
             row.names = FALSE, col.names = c("Persistent ID", "Downloads"))
```


### Institutional members
```{r member_projects}
# Can we detect institutional members from our routine earlier based on names?
deposit_members <- lapply(deposits$contacts, function(x, login) {
  names <- tibble(x=x) %>% hoist(x, "name") %>% .$name
  match <- str_which(login$full_name, paste(names, collapse="|"))[1]
  if (length(match)==0) return(NA)
  return(login$member[match])
}, login)
deposits <- tibble(member=deposit_members) %>% unnest(member) %>% bind_cols(deposits)

deposits %>% filter(!is.na(member)) %>% group_by(member, versionState) %>% 
  tally() %>%
  knitr::kable(caption = "Deposits associated with institution", 
               col.names = c("Institution", "Status", "Deposits"))

deposits %>% filter(!is.na(member) & as.Date(createdAt) > period_end) %>% 
  group_by(member, versionState) %>% 
  summarize(length(global_id)) %>%
  knitr::kable(caption=paste("Deposits since", period_end),
               col.names = c("Institution", "Status", "Deposits"))

projects %>% group_by(global_id=pid) %>% summarize(downloads=sum(count)) %>%
  inner_join(deposits) %>% select("name", "member", "downloads") %>%
  filter(!is.na(member)) %>% arrange(desc(downloads)) %>%
  group_by(member) %>% slice_head(n=3) %>%
  knitr::kable(caption="Top 3 downloads from each member institution",
               col.names = c("Deposit", "Member institution", "Downloads"))
```

## Make Data Count
```{r views}
views_total <- content(GET(paste0(dataverse, '/makeDataCount/viewsTotal/monthly')))
views_total <- bind_rows(views_total$data)
names(views_total) <- c("date", "total")

views_unique <- content(GET(paste0(dataverse, '/makeDataCount/viewsUnique/monthly')))
views_unique <- bind_rows(views_unique$data)
names(views_unique) <- c("date", "unique")

both <- merge(views_total, views_unique, all=TRUE, by="date")
both_ts <- ts(both[, -1], frequency=12, 
              start=strsplit(both$date[1], split="-")[[1]])
autoplot(both_ts, facets = FALSE) + labs(title="Views", color="Statistic") 
```

```{r downloads}
downloads_total <- content(GET(paste0(dataverse, '/makeDataCount/downloadsTotal/monthly')))
downloads_total <- bind_rows(downloads_total$data)
names(downloads_total) <- c("date", "total")

downloads_unique <- content(GET(paste0(dataverse, '/makeDataCount/downloadsUnique/monthly')))
downloads_unique <- do.call(rbind.data.frame, downloads_unique$data)
names(downloads_unique) <- c("date", "unique")

both <- merge(downloads_total, downloads_unique, all=TRUE, by="date")
both_ts <- ts(both[, -1], frequency=12, 
              start=strsplit(both$date[1], split="-")[[1]])
autoplot(both_ts, facets = FALSE) +  labs(title="Downloads", color="Statistic") 
```

